skeleton

### 2022-12-02 inst/skeleton.Rmd
typos

We can *try* to automate the above cycle as follows.  The first step
is to convert `skeleton_table` to a numeric matrix, which proves to be
a very awkward process

NB: `data.matrix()` fucks things up:

@@

Above, we see that Grotheer apparently came 17th in run 4, clearly
some weird infelicity.  OK, moving swiftly on, can we coerce our
dataframe to a numeric matrix?  Yes, we can, but it's klunky:

@@

Taking Maier as an example (a mid-range competitor) we want to
identify competitors who beat Maier at least once, and competitors who
Maier beats at least once (these groups are not necessarily distinct)

```{r}
bm1 <- st[,1] < st["Maier",1]  # beats Maier in run 1
bm2 <- st[,2] < st["Maier",2]  # beats Maier in run 2
bm3 <- st[,3] < st["Maier",3]  # beats Maier in run 3
bm4 <- st[,4] < st["Maier",4]  # beats Maier in run 4
bm1 | bm2 | bm3 | bm4
st[bm1 | bm2 | bm3 | bm4,]
```

Above we see all competitors who beat `Maier` at least once.  Who is
the worst of these, as measured by his worst performance?  This would
be `Weston`, who ranked 14 in run 1, compared to `Maier` who ranked
15.  Idiomatically:

```{r}
jj <- st[bm1 | bm2 | bm3 | bm4,]
which(jj==max(jj),arr.ind=TRUE)
rownames(jj)[which(jj==max(jj),arr.ind=TRUE)[1]]
```

We can automate this process as follows:

```{r}
wf <- function(p){
 beats <- function(run){st[,run] < st[p,run]}
 jj <- st[apply(sapply(seq_len(ncol(st)),beats),1,any),]
 rownames(jj)[which(jj==max(jj),arr.ind=TRUE)[1]]
}
wf("Maier")
```

To check:

```{r}
wf("Zheng")
```

Above we see that `Dukurs_M` beats `Zheng` in run 7, but `Dukurs_M`
ranked only 13 in run 4.

```{r}
loop <- c("Grotheer")
for(i in 1:20){loop <- c(loop,wf(loop[length(loop)]))}
loop
```

Above we see an endless repetition of `Blaser` -> `Timmings` -> `Blaser`.

```{r}
loop <- c("Jungk")
for(i in 1:20){loop <- c(loop,wf(loop[length(loop)]))}
loop
```

```{r}
loop <- c("Wengang")
for(i in 1:20){loop <- c(loop,wf(loop[length(loop)]))}
loop
```

```{r}
forw <- function(start){
  loop <- start
  while(all(table(loop)==1)){ loop <- c(loop,wf(loop[length(loop)])) }
  return(loop)
}
forw("Wengang")
```

We can do the process in reverse:

```{r}
fw <- function(p){
 beatenby <- function(run){st[,run] > st[p,run]}
 jj <- st[apply(sapply(seq_len(ncol(st)),beatenby),1,any),]
 rownames(jj)[which(jj==min(jj),arr.ind=TRUE)[1]]
}
fw("Maier")
```

Above, we see that `Dukurs_M` is the best (ranking 3 in run 2) of
competitors who lose to `Maier` [11 vs 13 in run 4].

```{r}
back <- function(start){
  loop <- start
  while(all(table(loop)==1)){ loop <- c(loop,fw(loop[length(loop)])) }
  return(loop)
}
back("Wengang")
```

```{r}
gd <- graph(c(1,2, 2,3, 1,2, 2,4, 1,4, 5,5, 3,6))
plot(gd)
path <- function(a){c(t(cbind(a[-length(a)],a[-1])))}

gpath <- NULL
jjf <- sapply(rownames(st),forw)
jjb <- sapply(rownames(st),back)
for(i in seq_along(jjf)){gpath <- c(gpath,path(jjf[[i]]))}
for(i in seq_along(jjb)){gpath <- c(gpath,rev(path(jjb[[i]])))}
gpath <- substr(gpath,1,3)
plot(graph(gpath[1:20]))
plot(graph(gpath[1:40]))
plot(graph(gpath[1:80]))
```

@@

```{r,label=processskeleton,cache=TRUE}
skeleton <- ordertable2supp(skeleton_table)
skeleton_maxp <- maxp(skeleton)
skeleton_maxp
pie(skeleton_maxp)
```

@@

>>>>

We can *try* to automate the above cycle as follows.  The first step
is to convert `skeleton_table` to a numeric matrix, which proves to be
a very awkward process.  NB: `data.matrix()` buggers things up:

@@

Above, we see that Grotheer apparently came 17th in run 4, clearly
some weird infelicity.  OK, how to coerce our dataframe to a numeric
matrix?  Possible but _extremely_  klunky:

@@

Taking Maier as an example (a mid-range competitor) we want to
identify competitors who Maier beats at least once:

```{r}
bm1 <- st[,1] > st["Maier",1]  # lost to Maier in run 1
bm2 <- st[,2] > st["Maier",2]  # lost to Maier in run 2
bm3 <- st[,3] > st["Maier",3]  # lost to Maier in run 3
bm4 <- st[,4] > st["Maier",4]  # lsot to Maier in run 4
st[bm1 | bm2 | bm3 | bm4,]
```

Above we see a subset of rows of  `skeleton_table` corresponding to
competitors who lost to Maier.

```{r}
rownames(st[bm1 | bm2 | bm3 | bm4,])
jj <- cbind(won="Maier",lost=rownames(st[bm1 | bm2 | bm3 | bm4,]))
noquote(jj)
plot(graph(c(t(jj))))
```

Above we see a directed graph, the arrow means "beats", so for example
Maier beats Timmings (2 o'clock).  We can functionify the calculation
and define `beats()`:

```{r}
beats <- function(p){   # returns everyone whom competitor 'p' beats
  bm1 <- st[,1] > st[p,1]  
  bm2 <- st[,2] > st[p,2]  
  bm3 <- st[,3] > st[p,3] 
  bm4 <- st[,4] > st[p,4] 
  jj <- cbind(p,rownames(st[bm1 | bm2 | bm3 | bm4,]))
  return(c(t(jj)))
}
bb <- c()
for(i in rownames(st)){bb <- c(bb,beats(i))}
graph(bb)
plot(graph(bb))
```

Above we see a directed graph.  The aim was to detect dominated sets,
that is, a subset of competitors who are always beaten by everyone
outside the group.  But although there is no such group in
`skeleton_table`, this fact is not immmediately obvious from the
graph.  We can however consider just the top and bottom few
competitors and see if that makes a difference.

```{r}
(st2 <- st[-(4:19),])  # exclude a bunch of mid-range competitors
```

Above, it is pretty clear that there is a dominated set of everyone
except the top 3.

```{r}
bb <- c()
for(i in rownames(st2)){bb <- c(bb,beats(i))}
graph(bb)
plot(graph(bb),layout = layout.fruchterman.reingold)
```

Above we see that the existence of the dominated set is not obvious
without close scrutiny.  Observing that Grotheer, Jungk, and Wengang
dominate the other competitors, we can see that these three have
arrows pointing to each other within the group, but noone outside the
group has an arrow pointing to any member of the set.  But this is not
obvious from inspection.

@@

```{r,label=processskeleton,cache=TRUE}
skeleton <- ordertable2supp(skeleton_table)
skeleton_maxp <- maxp(skeleton)
skeleton_maxp
pie(skeleton_maxp)
dotchart(skeleton_maxp)
dotchart(log(skeleton_maxp))
```

@@

Not very strong.

```{r,label=dgdtert,cache=TRUE}
samep.test(skeleton,c("Grotheer","Jungk"))$p.value
```

```{r,label=dgdtkkkert,cache=TRUE}
p <- rownames(st)
pvals <- seq_along(p)
for(i in seq(from=2,to=length(p))){ pvals[i] <- samep.test(skeleton,c(p[1],p[i]))$p.value }
names(pvals) <- p
pvals
```

(NB the above chunk takes about an hour to run, even with `n=1`).

```{r}
plot(pvals,log='y',pch=16)
for(i in seq_along(pvals)){text(i,pvals[i],names(pvals)[i],pos=4,col='gray')}
```
